Compositional Certified Resource Bounds\ifshort{\vspace{-.15cm}}{}

### 2014-11-11 paper/main.tex
minor
new abstract

\begin{abstract}
This paper presents a new approach to quantitative
analysis for C programs.  The technique described uses
ideas from amortized analysis and abstract interpretation
to derive tight linear worst-case bounds that come with a
proof certificate.  Contrary to most available tools for
C, ours has a compositional design that provides the
programmer with bounds expressed in terms of meaningful
variables.  The compositionality allows to handle uniformly
nested sequenced loops and recursive sequenced function
calls.  The automatic analysis is implemented
as a tool that leverages state-of-the-art SMT solvers
LP solvers to automatically infer resource bounds on
a wide range of C programs.  It is the first automatic
amortized analysis that can derive bounds that depend
on negative integers and differences between numbers.
To prove its soundness and give a concrete shape to proof
certificates, we developed a quantitative Hoare logic
that we proved sound using the Coq proof assistant.  This
logic allows manual proofs of arbitrarily complex resource
bounds on C programs and interaction with the automation.
To prove the scalability and precision of this new tool,
we run it on multiple code snippets from the literature,
reproduce results from previous work, and analyze a
consequent part of a modern benchmark suite.

@@

$3.1|[y,z]| + 0.1|[0,y]|$
\\[.4\baselineskip]
      {\bf t08a}
    \end{center}
  \end{minipage}
%
\hfill
%
%
%
  \begin{minipage}[b]{\progwidth}
    \begin{center}
   \begin{lstlisting}
while (n<0) {
  ${\color{blue}%
  n{<}0; \,  P(n,y) \vdash }$
  n=n+1;
  ${\color{blue}%
  \dashv \cdot; \,  59 {+} P(n,y) \vdash }$
  y=y+1000;
  ${\color{blue}%
  \dashv \cdot; \, 9 {+} P(n,y) \vdash }$
  while (y>=100 && *){
    ${\color{blue}%
    y{>}99; \, 9 {+} P(n,y) \vdash }$
    y=y-100;
    ${\color{blue}%
    \dashv \cdot; \, 14 {+} P(n,y) \vdash }$
    tick(5);
    ${\color{blue}%
    \dashv \cdot; \, 9 {+} P(n,y) }$
  }
  ${\color{blue}%
  \dashv \cdot; \, 9 {+} P(n,y) \vdash }$
  tick(9);
  ${\color{blue}%
  \dashv \cdot; \, P(n,y) }$
}
   \end{lstlisting}
$59|[n,0]| {+} 0.05|[0,y]|$
\\[.4\baselineskip]
      {\bf t27}
    \end{center}
  \end{minipage}
\vspace{1ex}
\caption{Derivations of bounds on the number of ticks for challenging
  examples.  Examples \emph{speed\_1} and \emph{speed\_2}
  (from~\cite{GulwaniMC09}) use \emph{tricky iteration patterns},
  \emph{t08a} contains \emph{sequential loops} so that the iterations
  of the second loop depend on the first, and \emph{t27} contains
  interacting \emph{nested loops}. In the potential functions, we only
  mention the non-zero terms and in the logical context $\Gamma$ we
  only mention assertions that we actually use in the reasoning. In
  Example \emph{t27}, we use the abbreviation $P(n,y) := 59|[n,0]| {+}
  0.05|[0,y]|$.}
  \label{fig:ex_list}
\end{figure*}

>>>>

\begin{abstract}
  This paper presents a new approach for automatically deriving
  worst-case resource bounds for C programs.
  %
  The described technique combines ideas from amortized analysis and
  abstract interpretation in a unified framework to address four
  challenges for state-of-the-art techniques: compositionality, user
  interaction, generation of proof certificates, and scalability.
  %
  \emph{Compositionality} is achieved by incorporating the potential
  method of amortized analysis.  It enables the derivation of global
  whole-program bounds with local derivation rules by naturally
  tracking size changes of variables in sequenced loops and function
  calls.  The resource consumption of functions is described abstractly
  and a function call can be analyzed without access to the function
  body.
  %\
  \emph{User interaction} is supported with a new mechanism that
  clearly separates qualitative and quantitative verification. A user
  can guide the analysis to derive complex non-linear bounds by using
  auxiliary variables and assertions.  The assertions are
  separately proved using established qualitative techniques such as
  abstract interpretation or Hoare logic.
  %
  \emph{Proof certificates} are automatically generated from the local
  derivation rules. A soundness proof of the derivation system with
  respect to a formal cost semantics guarantees the validity of the
  certificates.
  %
  \emph{Scalability} is attained by an efficient reduction of
  bound inference to a linear optimization problem that can be solved
  by off-the-shelf LP solvers.
  %
  The analysis framework is implemented in the publically-available
  tool $\toolname$. An experimental evaluation demonstrates the
  advantages of the new technique with a comparison of $\toolname$ with
  existing tools on challenging micro benchmarks and the analysis of
  more than XXX lines of C code from the XXX benchmark suite.

% Contrary to most available tools for
% C, ours has a compositional design that provides the
% programmer with bounds expressed in terms of meaningful
% variables.  The compositionality allows to handle uniformly
% nested sequenced loops and recursive sequenced function
% calls.  The automatic analysis is implemented
% as a tool that leverages state-of-the-art SMT solvers
% LP solvers to automatically infer resource bounds on
% a wide range of C programs.  It is the first automatic
% amortized analysis that can derive bounds that depend
% on negative integers and differences between numbers.
% To prove its soundness and give a concrete shape to proof
% certificates, we developed a quantitative Hoare logic
% that we proved sound using the Coq proof assistant.  This
% logic allows manual proofs of arbitrarily complex resource
% bounds on C programs and interaction with the automation.
% To prove the scalability and precision of this new tool,
% we run it on multiple code snippets from the literature,
% reproduce results from previous work, and analyze a
% consequent part of a modern benchmark suite.

@@

$3.1|[y,z]| + 0.1|[0,y]|$
\\[.4\baselineskip]
      {\bf t08a}
    \end{center}
  \end{minipage}
%
\hfill
%
%
%
  \begin{minipage}[b]{\progwidth}
    \begin{center}
   \begin{lstlisting}
while (n<0) {
  ${\color{blue}%
  n{<}0; \,  P(n,y) \vdash }$
  n=n+1;
  ${\color{blue}%
  \dashv \cdot; \,  59 {+} P(n,y) \vdash }$
  y=y+1000;
  ${\color{blue}%
  \dashv \cdot; \, 9 {+} P(n,y) \vdash }$
  while (y>=100 && *){
    ${\color{blue}%
    y{>}99; \, 9 {+} P(n,y) \vdash }$
    y=y-100;
    ${\color{blue}%
    \dashv \cdot; \, 14 {+} P(n,y) \vdash }$
    tick(5);
  } ${\color{blue}%
  \dashv \cdot; \, 9 {+} P(n,y) \vdash }$
  tick(9);
  ${\color{blue}%
  \dashv \cdot; \, P(n,y) }$
}
   \end{lstlisting}
$59|[n,0]| {+} 0.05|[0,y]|$
\\[.4\baselineskip]
      {\bf t27}
    \end{center}
  \end{minipage}
\vspace{1ex}
\caption{Derivations of bounds on the number of ticks for challenging
  examples.  Examples \emph{speed\_1} and \emph{speed\_2}
  (from~\cite{GulwaniMC09}) use \emph{tricky iteration patterns},
  \emph{t08a} contains \emph{sequential loops} so that the iterations
  of the second loop depend on the first, and \emph{t27} contains
  interacting \emph{nested loops}. In the potential functions, we only
  mention the non-zero terms and in the logical context $\Gamma$ we
  only mention assertions that we actually use in the reasoning. In
  Example \emph{t27}, we use the abbreviation $P(n,y) := 59|[n,0]| {+}
  0.05|[0,y]|$.}
  \label{fig:ex_list}
\end{figure*}
